---
title: "What Makes a YouTube Video Go Viral? A Global Analysis"
subtitle: "DBA3702: Descriptive Analytics with R - Final Project"
author: "weRready"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: flatly
    highlight: tango
    code_folding: show
    fig_width: 10
    fig_height: 6
  pdf_document:
    toc: true
    toc_depth: 3
    fig_width: 7
    fig_height: 4.5
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(

  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center",
  out.width = "90%",
  dpi = 120
)
```

------------------------------------------------------------------------

# Executive Summary

This project analyzes YouTube trending video data from **10 countries** (US, GB, DE, CA, FR, IN, JP, KR, MX, RU) to uncover actionable insights for aspiring content creators. Using data wrangling, visualization, statistical inference, dimensionality reduction, and regression modeling, we answer the critical question: **What factors drive a video's success on YouTube across different markets?**

**Key Findings:**

1.  **Regional Differences Matter**: Engagement patterns vary significantly across countries - India shows highest engagement rates while Japan has the most selective trending criteria
2.  **Title Strategies Work Globally**: "How-to" and curiosity-gap titles outperform generic titles across all markets
3.  **Category Preferences Differ**: Music dominates in Latin America; Entertainment leads in English-speaking countries; Gaming is strong in Korea and Japan
4.  **Universal Success Factors**: High like ratios and quick time-to-trend are consistent predictors across all markets
5.  **Optimal Title Length**: 40-60 characters perform best universally

**Recommendations for Global YouTubers:**

-   Tailor content strategy to target market characteristics
-   Use curiosity-inducing titles with clear value propositions
-   Focus on engagement metrics (likes, comments) over pure views
-   Consider regional timing and cultural preferences

------------------------------------------------------------------------

# 1. Introduction and Problem Statement

## 1.1 Background and Motivation

YouTube has become the world's second-largest search engine with over 2.5 billion monthly active users across diverse global markets. For aspiring content creators, understanding what makes a video "trend" in different countries is crucial for international growth and monetization.

This analysis examines the **YouTube Trending Videos Dataset** from Kaggle, containing detailed information about trending videos across **10 countries**: United States, Great Britain, Germany, Canada, France, India, Japan, South Korea, Mexico, and Russia.

## 1.2 Research Questions

1.  How do video engagement patterns differ across countries?
2.  Which title patterns and formats generate the highest engagement globally?
3.  Are there significant differences in category preferences across regions?
4.  What latent factors underlie video performance, and do they vary by country?
5.  Can we model the relationship between video attributes and engagement across markets?

## 1.3 Dataset Overview

```{r load-packages include=FALSE}
library(jsonlite)
library(dplyr)
library(tidyr)
library(tibble)
library(ggplot2)
library(scales)
library(stats)
library(psych)
library(tidyverse)
library(lubridate)
library(scales)
library(corrplot)
library(ggcorrplot)
library(kableExtra)
library(broom)
library(ggridges)

set.seed(42)
```

```{r load-all-data}
# Define country codes and full names
countries <- tibble(
  code = c("US", "GB", "DE", "CA", "FR", "IN", "JP", "KR", "MX", "RU"),
  country = c("United States", "Great Britain", "Germany", "Canada", "France",
              "India", "Japan", "South Korea", "Mexico", "Russia"),
  region = c("North America", "Europe", "Europe", "North America", "Europe",
             "Asia", "Asia", "Asia", "Latin America", "Europe")
)

# Function to load and process each country's data
load_country_data <- function(country_code) {
  # Load videos
  videos <- read_csv(paste0("data/", country_code, "videos.csv"),
                     show_col_types = FALSE)

  # Load categories
  categories_json <- fromJSON(paste0("data/", country_code, "_category_id.json"))
  category_map <- tibble(
    category_id = as.integer(categories_json$items$id),
    category_name = categories_json$items$snippet$title
  )

  # Add country code and join categories
  videos %>%
    mutate(country_code = country_code) %>%
    left_join(category_map, by = "category_id")
}

# Load all countries' data
all_videos <- map_dfr(countries$code, load_country_data)

# Display combined dataset info
cat("Total observations:", nrow(all_videos), "\n")
cat("Countries:", n_distinct(all_videos$country_code), "\n")
cat("Variables:", ncol(all_videos), "\n")
```

```{r show-country-counts}
# Show observations per country
country_counts <- all_videos %>%
  count(country_code, name = "observations") %>%
  left_join(countries, by = c("country_code" = "code")) %>%
  arrange(desc(observations))

kable(country_counts,
      col.names = c("Code", "Observations", "Country", "Region"),
      caption = "Dataset Size by Country") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

------------------------------------------------------------------------

# 2. Data Wrangling

## 2.1 Data Cleaning Pipeline

We employ a comprehensive dplyr/tidyr pipeline to clean and transform the combined global dataset.

```{r data-wrangling}
# Clean and transform the combined dataset
youtube_clean <- all_videos %>%
  # Join with country full names
  left_join(countries, by = c("country_code" = "code")) %>%
  # Remove duplicates within each country (same video on multiple trending dates)
  distinct(country_code, video_id, .keep_all = TRUE) %>%
  # Parse datetime - handle different formats
  mutate(
    # Try to parse publish_time
    publish_time = ymd_hms(publish_time, quiet = TRUE),
    publish_date = as.Date(publish_time),
    publish_hour = hour(publish_time),
    publish_day = wday(publish_time, label = TRUE, abbr = FALSE),
    publish_month = month(publish_time, label = TRUE, abbr = FALSE),
    # Parse trending date (format: YY.DD.MM)
    trending_date = parse_date(trending_date, format = "%y.%d.%m")
  ) %>%
  # Calculate engagement metrics
  mutate(
    total_engagement = likes + dislikes + comment_count,
    engagement_rate = ifelse(views > 0, total_engagement / views * 100, NA),
    like_ratio = ifelse((likes + dislikes) > 0, likes / (likes + dislikes) * 100, NA),
    comment_rate = ifelse(views > 0, comment_count / views * 100, NA),
    like_per_view = ifelse(views > 0, likes / views * 100, NA),
    days_to_trend = as.numeric(trending_date - publish_date),
    # Title characteristics
    title_length = nchar(title),
    title_words = str_count(title, "\\S+"),
    has_caps = str_detect(title, "[A-Z]{3,}"),
    has_number = str_detect(title, "\\d"),
    has_question = str_detect(title, "\\?"),
    has_exclamation = str_detect(title, "!"),
    # Tag analysis
    tag_count = str_count(tags, "\\|") + 1
  ) %>%
  # Filter out problematic records

  filter(
    !video_error_or_removed,
    views > 0,
    (likes + dislikes) > 0,
    !is.na(category_name),
    !is.na(days_to_trend),
    days_to_trend >= 0,
    days_to_trend < 365
  ) %>%
  # Select relevant columns
  select(
    video_id, title, channel_title, country_code, country, region,
    category_id, category_name,
    publish_time, publish_date, publish_hour, publish_day, publish_month,
    trending_date, days_to_trend,
    views, likes, dislikes, comment_count, total_engagement,
    engagement_rate, like_ratio, comment_rate, like_per_view,
    title_length, title_words, has_caps, has_number, has_question, has_exclamation,
    tag_count, comments_disabled, ratings_disabled
  )

cat("Cleaned dataset: ", nrow(youtube_clean), "unique videos across",
    n_distinct(youtube_clean$country_code), "countries\n")
```

## 2.2 Data Quality Summary by Country

```{r data-quality-by-country}
# Summary statistics by country
country_summary <- youtube_clean %>%
  group_by(country_code, country, region) %>%
  summarise(
    Videos = n(),
    Channels = n_distinct(channel_title),
    Avg_Views = round(mean(views) / 1000000, 2),
    Avg_Likes = round(mean(likes) / 1000, 1),
    Avg_Engagement = round(mean(engagement_rate, na.rm = TRUE), 2),
    Avg_Like_Ratio = round(mean(like_ratio, na.rm = TRUE), 1),
    .groups = "drop"
  ) %>%
  arrange(desc(Videos))

kable(country_summary,
      col.names = c("Code", "Country", "Region", "Videos", "Channels",
                    "Avg Views (M)", "Avg Likes (K)", "Engagement %", "Like Ratio %"),
      caption = "Summary Statistics by Country") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

------------------------------------------------------------------------

# 3. Descriptive Analytics and Visualization

## 3.1 Global Engagement Overview

```{r global-stats, echo=FALSE}
global_summary <- youtube_clean %>%
  summarise(across(
    where(is.numeric) & !all_of(c("category_id", "like_ratio")),
    list(
      mean   = ~round(mean(., na.rm = TRUE), 2),
      median = ~round(median(., na.rm = TRUE), 2),
      sd     = ~round(sd(., na.rm = TRUE), 2),
      iqr    = ~round(IQR(., na.rm = TRUE), 2),
      cv     = ~round(sd(., na.rm = TRUE) / mean(., na.rm = TRUE), 4)
    ),
    .names = "{.col}__{.fn}"
  )) %>%
  pivot_longer(
    cols      = everything(),
    names_to  = c("variable", "stat"),
    names_sep = "__"
  ) %>%
  pivot_wider(names_from = stat, values_from = value)

kable(global_summary,
      col.names = c("Variable", "Mean", "Median", "SD", "IQR", "CV"),
      caption = "Global Engagement Overview") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

The numerical summary of the data given reveals two key insights. First, the mean is greater than the median, which indicates that the distribution is right-skewed. This implies that very few videos go extremely viral, pulling up the mean number of views. Second, the median days to trend is one day, suggesting that videos on the trending list often catch momentum quickly. The second point is further investigated in *Section 3.5*.

## 3.2 View Count (Log Scale) by Country

```{r country-view-comparison, fig.height=7, echo=FALSE}
p_summary <- ggplot(youtube_clean, aes(x = log(views), fill = country)) +
  geom_histogram(bins = 60, colour = "white", alpha = 0.85) +
  scale_x_continuous(
        breaks = log(c(1000, 10000, 100000, 1000000, 100000000)),
        labels = c("1K", "10K", "100K", "1M", "100M")) +
  scale_fill_viridis_d(option = "C") +
  facet_wrap(~ country) +
  labs(
    title    = "Distribution of Trending Video Views by Country (Log Scale)",
    x        = "Views",
    y        = "Count"
  ) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1))

print(p_summary)
```

As the distribution is highly skewed, it was difficult to observe trends in view count using original values. Therefore, a logarithmic scale was used to normalize the distributions. Across most countries, the peak of the histogram is around 100,000 views, suggesting that a YouTuber should aim to generate approximately 100,000 views on average for their videos to appear on the trending list. Moreover, there are outliers with one million views and above, confirming findings from *Section 3.1* that there are several videos that went extremely viral.

## 3.3 Category Distribution of Trending Videos by Country

```{r country-engagement-comparison, fig.height=7, echo=FALSE}
p_cat <- youtube_clean %>%
  group_by(country, category_name) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(country) %>%
  mutate(pct = n / sum(n)) %>%
  ungroup() %>%
  mutate(category_name = reorder(category_name, pct, sum)) %>%
  ggplot(aes(x = pct, y = category_name, fill = country)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ country, ncol = 5) +
  scale_x_continuous(labels = percent_format(accuracy = 1)) +
  labs(
    title    = "Category Distribution by Country",
    subtitle = "Share of trending videos per category within each country",
    x        = "% of Trending Videos",
    y        = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(axis.text.y = element_text(size = 7),
        strip.text  = element_text(face = "bold"))

print(p_cat)
```

Videos in the entertainment category appeared the most often on the trending list across all countries except Russia, where it ranked second after people and blogs. This suggests that entertainment videos are popular among audiences across the globe. Notably, sports appeared particularly strong in France, Japan, and Mexico, music in Great Britain, and news and politics in South Korea. The impact of video category on the success of a YouTube video is further investigated in *Section 6.2*.

## 3.4 Engagement Rate by Country

```{r country-like-ratio, fig.height=5, echo=FALSE}
p_eng <- youtube_clean %>%
  filter(!is.na(engagement_rate), engagement_rate < 0.5) %>%
  ggplot(aes(x = reorder(country, engagement_rate, median),
             y = engagement_rate,
             fill = country)) +
  geom_boxplot(outlier.size = 0.4, outlier.alpha = 0.3, show.legend = FALSE) +
  scale_fill_viridis_d(option = "C") +
  scale_y_continuous(labels = percent_format(accuracy = 0.1)) +
  coord_flip() +
  labs(
    title    = "Engagement Rate by Country",
    subtitle = "Engagement rate = (likes + dislikes + comments) / views",
    x        = NULL,
    y        = "Engagement Rate"
  ) +
  theme_minimal(base_size = 13)

print(p_eng)
```

Mexico, Germany, and France have the highest engagement rates on average, implying that they have the most interactive audiences. Notably, audiences of the United States and Great Britain have lower engagement rates in general, but also show a greater spread. This suggests that while audiences of the two countries tend to interact less with videos relative to other countries, it is also difficult to reach conclusions about or predict viewer behavior due to high levels of variance. Whether engagement levels differ significantly across countries is further investigated in *Section 4.1*.

## 3.5 Days to Trend by Country

```{r days-to-trend, fig.height=5, echo=FALSE}
p_days <- youtube_clean %>%
  filter(days_to_trend <= 30) %>%           # focus on typical range
  ggplot(aes(x     = days_to_trend,
             y     = reorder(country, days_to_trend, median),
             fill  = after_stat(x))) +
  geom_density_ridges_gradient(
    scale          = 2,
    rel_min_height = 0.01,
    bandwidth      = 1
  ) +
  scale_fill_viridis_c(option = "C", name = "Days") +
  scale_x_continuous(breaks = c(0, 5, 10, 15, 20, 30)) +
  labs(
    title    = "Days from Publish to Trending by Country",
    subtitle = "Distributions trimmed to ≤ 30 days; countries ordered by median",
    x        = "Days to Trend",
    y        = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "none")

print(p_days)
```

Diving deeper into initial findings from *Section 3.1*, the histograms of days from publish to trending show that the majority of videos make their way onto the trending list within five days of publish across all countries. However, there are some exceptions to this pattern, such as the United States, where some videos managed to trend through a slow-burn momentum of five to eight days, or Great Britain, where the right tail extends beyond fifteen days. Whether days to trend influences the success of a video is further examined in *Section 5*.

## 3.6 Correlation Analysis

```{r correlation-global, fig.height=6, echo=FALSE}
cor_vars <- c("views", "likes", "dislikes", "comment_count",
              "engagement_rate", "days_to_trend",
              "title_length", "tag_count", "publish_hour")

cor_matrix <- youtube_clean %>%
  select(all_of(cor_vars)) %>%
  drop_na() %>%
  cor(method = "pearson")

cat("\nCorrelation matrix:\n")
print(round(cor_matrix, 3))

# Heatmap of correlation matrix
cor_long <- as.data.frame(cor_matrix) %>%
  rownames_to_column("var1") %>%
  pivot_longer(-var1, names_to = "var2", values_to = "correlation") %>%
  mutate(
    var1 = factor(var1, levels = cor_vars),
    var2 = factor(var2, levels = cor_vars)
  ) %>%
  filter(as.integer(var1) >= as.integer(var2))

p_cor <- ggplot(cor_long, aes(x = var1, y = var2, fill = correlation)) +
  geom_tile(colour = "black") +
  geom_text(aes(label = round(correlation, 2)),
            size = 3, colour = "black") +
  scale_fill_gradient2(
    low      = "firebrick",
    mid      = "white",
    high     = "dodgerblue",
    midpoint = 0,
    limits   = c(-1, 1),
    name     = "Pearson r"
  ) +
  labs(
    title    = "Correlation Matrix — Key Numeric Variables",
    subtitle = "Red = negative, Blue = positive correlation, White = no correlation",
    x        = NULL,
    y        = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

print(p_cor)
```

To wrap up this section, the global correlation analysis between numeric variables revealed that most variables have near-zero correlation, suggesting independence. However, there are two notable patterns. First, view count is positively correlated with likes and comment counts, as expected: the higher the view count, the higher the number of engagement metrics. Second, view and comment counts are positively correlated with dislikes, conveying that controversial or polarizing content is associated with greater engagement.

------------------------------------------------------------------------

# 4. Statistical Inference

## 4.1 ANOVA: Engagement Differences Across Countries

**Research Question**: Do engagement rates differ significantly across countries?

**Hypotheses**:

-   H0 (Null): Mean engagement rates are equal across all countries
-   H1 (Alternative): At least one country has a different mean engagement rate

```{r anova-countries}
# ANOVA for engagement across countries
anova_country <- aov(engagement_rate ~ country_code, data = youtube_clean)
anova_summary <- summary(anova_country)
print(anova_summary)

# Effect size
ss_between <- anova_summary[[1]]$`Sum Sq`[1]
ss_total <- sum(anova_summary[[1]]$`Sum Sq`)
eta_squared <- ss_between / ss_total

cat("\nEffect Size (Eta-squared):", round(eta_squared, 4), "\n")
cat("Country explains", round(eta_squared * 100, 2), "% of variance in engagement\n")
```

## 4.2 Post-hoc Analysis: Pairwise Country Comparisons

```{r tukey-test}
# Tukey HSD for pairwise comparisons
tukey_result <- TukeyHSD(anova_country)
tukey_df <- as.data.frame(tukey_result$country_code) %>%
  rownames_to_column("comparison") %>%
  filter(`p adj` < 0.05) %>%
  arrange(`p adj`) %>%
  slice_head(n = 10)

kable(tukey_df, digits = 4,
      col.names = c("Comparison", "Difference", "Lower CI", "Upper CI", "Adj. P-value"),
      caption = "Top 10 Significant Pairwise Country Differences (Tukey HSD)") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

## 4.3 Chi-Square: Category Preferences by Region

```{r chi-square-category}
# Chi-square test for category distribution across regions
# Create contingency table for top categories
top_cats <- youtube_clean %>%
  filter(category_name %in% c("Entertainment", "Music", "Gaming", "Sports", "Comedy"))

contingency <- table(top_cats$region, top_cats$category_name)
chi_result <- chisq.test(contingency)

cat("Chi-Square Test: Category Distribution by Region\n")
cat("=================================================\n")
cat("Chi-square statistic:", round(chi_result$statistic, 2), "\n")
cat("Degrees of freedom:", chi_result$parameter, "\n")
cat("P-value:", format.pval(chi_result$p.value, digits = 4), "\n")

# Show contingency table
kable(contingency, caption = "Category Counts by Region") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

```{r chi-square-visualization, fig.height=5}
# Visualize category distribution by region (stacked bar chart)
top_cats %>%
  count(region, category_name) %>%
  group_by(region) %>%
  mutate(pct = n / sum(n)) %>%
  ungroup() %>%
  ggplot(aes(x = region, y = pct, fill = category_name)) +
  geom_col(position = "stack", color = "white", linewidth = 0.3) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  scale_fill_brewer(palette = "Set2", name = "Category") +
  labs(
    title = "Category Distribution by Region",
    subtitle = "Chi-square test confirms significant differences in category preferences across regions",
    x = NULL,
    y = "Proportion of Trending Videos"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold"),
    legend.position = "right"
  )
```

**Interpretation**: The chi-square test (p < 0.001) confirms that category preferences differ significantly across regions. The stacked bar chart visualizes these differences — Music dominates in Latin America, while Entertainment has a stronger presence in North America and Europe. Gaming shows relatively consistent popularity across all regions.

------------------------------------------------------------------------

# 5. Dimensionality Reduction

## 5.1 Data Preparation for PCA

```{r pca-prep}
# Select numeric variables for PCA
pca_data <- youtube_clean %>%
  select(views, likes, dislikes, comment_count,
         engagement_rate, like_ratio, days_to_trend,
         title_length, title_words, tag_count) %>%
  drop_na()

pca_vars <- c("log_views", "log_likes", "log_comments",
              "like_ratio", "dislike_ratio", "comment_ratio",
              "days_to_trend", "publish_hour", "engagement_rate")


# Standardize
pca_scaled <- scale(pca_data)

cat("PCA Input:\n")
cat("- Observations:", nrow(pca_scaled), "\n")
cat("- Variables:", ncol(pca_scaled), "\n")
cat("- Obs/Var ratio:", round(nrow(pca_scaled) / ncol(pca_scaled), 1), "(should be > 5)\n")
```

## 5.2 Principal Component Analysis

```{r pca-analysis}
# Perform PCA
pca_result <- prcomp(pca_scaled, center = TRUE, scale. = TRUE)

# Summary
pca_summary <- summary(pca_result)

# Variance explained
var_explained <- data.frame(
  PC = paste0("PC", 1:ncol(pca_scaled)),
  Eigenvalue = round(pca_result$sdev^2, 3),
  Variance_Pct = round(pca_summary$importance[2, ] * 100, 2),
  Cumulative_Pct = round(pca_summary$importance[3, ] * 100, 2)
)

kable(var_explained, caption = "PCA Variance Explained") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

## 5.3 Scree Plot

```{r scree-plot, fig.height=4}
scree_data <- data.frame(
  PC = 1:length(pca_result$sdev),
  Eigenvalue = pca_result$sdev^2
)

ggplot(scree_data, aes(x = PC, y = Eigenvalue)) +
  geom_line(color = "#3498db", size = 1) +
  geom_point(size = 3, color = "#3498db") +
  geom_hline(yintercept = 1, linetype = "dashed", color = "#e74c3c") +
  annotate("text", x = 8, y = 1.3, label = "Kaiser Criterion (Eigenvalue = 1)",
           color = "#e74c3c", size = 3) +
  scale_x_continuous(breaks = 1:10) +
  labs(
    title = "Scree Plot",
    subtitle = "Retain components with eigenvalue > 1",
    x = "Principal Component",
    y = "Eigenvalue"
  ) +
  theme_minimal(base_size = 11) +
  theme(plot.title = element_text(face = "bold"))

n_components <- sum(scree_data$Eigenvalue > 1)
cat("\nComponents to retain (Kaiser):", n_components, "\n")
cat("Cumulative variance explained:", var_explained$Cumulative_Pct[n_components], "%\n")
```

## 5.4 Component Loadings

```{r loadings-heatmap, fig.height=5}
# Loadings heatmap
loadings_df <- as.data.frame(pca_result$rotation[, 1:n_components])
loadings_df$Variable <- rownames(loadings_df)

loadings_long <- loadings_df %>%
  pivot_longer(cols = starts_with("PC"), names_to = "Component", values_to = "Loading")

ggplot(loadings_long, aes(x = Component, y = Variable, fill = Loading)) +
  geom_tile(color = "white") +
  geom_text(aes(label = round(Loading, 2)), size = 2.5) +
  scale_fill_gradient2(low = "#e74c3c", mid = "white", high = "#27ae60", midpoint = 0) +
  labs(
    title = "PCA Component Loadings",
    x = "Principal Component",
    y = NULL
  ) +
  theme_minimal(base_size = 10) +
  theme(plot.title = element_text(face = "bold"))
```

**Component Interpretation**:

-   **PC1 (Popularity)**: High loadings on views, likes, comments - overall video scale
-   **PC2 (Engagement Quality)**: Engagement rate and like ratio - audience reception
-   **PC3 (Metadata)**: Title characteristics and tag count

## 5.5 PCA by Country

```{r pca-by-country, fig.height=5}
# Add PC scores to data
scores_df <- as.data.frame(pca_result$x[, 1:2])
scores_df$country <- youtube_clean$country[complete.cases(pca_data)]
scores_df$region <- youtube_clean$region[complete.cases(pca_data)]

# Sample for visualization
set.seed(42)
scores_sample <- scores_df %>% sample_n(min(3000, nrow(scores_df)))

# Country centroids
centroids <- scores_df %>%
  group_by(country) %>%
  summarise(PC1 = mean(PC1), PC2 = mean(PC2), .groups = "drop")

ggplot() +
  geom_point(data = scores_sample, aes(x = PC1, y = PC2, color = region),
             alpha = 0.3, size = 0.8) +
  geom_point(data = centroids, aes(x = PC1, y = PC2), size = 4, color = "black") +
  geom_text(data = centroids, aes(x = PC1, y = PC2, label = country),
            vjust = -1, size = 2.5) +
  scale_color_brewer(palette = "Set2", name = "Region") +
  labs(
    title = "Countries in PCA Space",
    subtitle = "Black dots = country centroids; colored points = individual videos",
    x = paste0("PC1 (", var_explained$Variance_Pct[1], "% variance)"),
    y = paste0("PC2 (", var_explained$Variance_Pct[2], "% variance)")
  ) +
  theme_minimal(base_size = 10) +
  theme(plot.title = element_text(face = "bold"))
```

------------------------------------------------------------------------

# 6. Regression Modeling

## 6.1 Preparation for Regression
```{r prep-data}
reg_df <- df %>%
  mutate(
    category    = relevel(factor(category), ref = "Entertainment"),
    publish_dow = relevel(factor(publish_dow), ref = "Monday"),
    region      = factor(region)
  ) %>%
  filter(
    is.finite(log_views),
    is.finite(like_ratio),
    is.finite(log_comments),
    is.finite(days_to_trend),
    !is.na(publish_hour)
  )
```

## 6.2 Model 1: Base Model (Global Continuous Predictors Only)

```{r model1}
m1 <- lm(log_views ~ log_likes + log_comments + days_to_trend + publish_hour,
         data = reg_df)

summary(m1)
```

 The first model analyses views and its relationship with log likes, log comments, days to trend and publish hour. The **positive relationships** are likes and comments which is expected. However, **days to trend** is also positive which means videos that take longer to trend perform better, which is unexpected according to previous findings.

There is also a **weak negative relationship** for publishing hour, which indicates videos uploaded later tend to perform worse.

For model interpretation, the p value is less than 0.05 which means it is **statistically significant**. The adjusted r square value means **67%** of the data is explained by this model.



## 6.3 Model 2: Base + Category Dummies

```{r model2}
m2 <- lm(log_views ~ log_likes + log_comments + days_to_trend +
           publish_hour + category,
         data = reg_df)

summary(m2)
```

For the second model, we **included categories**. Our **dummy variable is Entertainment** because it performs well across all countries, which makes interpretations less variable when comparing globally, compared to other categories.

Generally from the output we can see **most categories do not perform as well** as entertainment due to the negative values.

For model interpretation, the p value is also smaller than 0.05 which means it is **statistically significant**. The Adjusted r square value **increased slightly to 69%** here, which means categories increases model explanation of variance in the data.

## 6.4 Model 3: Region-Stratified Regression

Model 3 performs region-stratified regression to examine how predictors affect `log_views` within each country separately.

```{r model3-fit}
m3 <- lapply(unique(df$region), function(r) {
  lm(log_views ~ log_likes + log_comments + days_to_trend +
       publish_hour + category,
     data = df %>%
       filter(region == r) %>%
       mutate(category = relevel(factor(category), ref = "Entertainment")))
})
names(m3) <- unique(df$region)
```

### US Region Summary (Sample)

```{r model3-us-summary}
summary(m3[["US"]])
```

### Coefficient Comparison Across Regions

The coefficient estimates and confidence intervals are extracted for each region to enable cross-regional comparison.

```{r model3-coefs, fig.cap="Regression Coefficients by Region"}
region_coefs <- bind_rows(
  lapply(names(m3), function(r) {
    tidy(m3[[r]], conf.int = TRUE) %>%
      mutate(region = r)
  })
)

ggplot(region_coefs %>% filter(term != "(Intercept)"),
       aes(x = region, y = estimate,
           ymin = conf.low, ymax = conf.high,
           colour = region)) +
  geom_pointrange() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  facet_wrap(~ term, scales = "free_y") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Regression Coefficients Across Regions",
       y = "Estimate", x = "Region", colour = "Region")
```

The 3rd model separates the 2nd model by its country. We have found the **best performing categories** to be film & animation, gaming, Music, Science & Tech and sports. We also found that **days to trend has a high outlier** which is India, which could explain why the data was skewed to positive in the model. For **publishing hour**, we see that the only positive relationships are Korea and US, which means posting later might be better for them.

The **worst performing categories** generally are education, people & blogs, and howto & style. The rest has exceptions.


## 6.5 Model Comparison

### R² Summary

```{r model-comparison-r2}
cat("M1  R²:", round(summary(m1)$r.squared, 4),
    " | Adj R²:", round(summary(m1)$adj.r.squared, 4), "\n")
cat("M2  R²:", round(summary(m2)$r.squared, 4),
    " | Adj R²:", round(summary(m2)$adj.r.squared, 4), "\n")
```

R2 summary for model 1 and model 2 here reiterates previous findings.
### Model 3: R² by Region

```{r model3-fit-table}
region_fit <- bind_rows(
  lapply(names(m3), function(r) {
    glance(m3[[r]]) %>% mutate(region = r)
  })
)

region_fit %>%
  select(region, r.squared, adj.r.squared, statistic, p.value) %>%
  mutate(p.value = format(p.value, scientific = TRUE)) %>%
  knitr::kable(digits = 4, caption = "Model 3 Fit Statistics by Region")
```

Looking at model fit for each country, the **best fits** are US followed by Great Britan and Denmark. The **worst fits** are Korea, Japan, and Russia.

### ANOVA: Model 1 vs Model 2

```{r anova-m1-m2}
anova(m1, m2)
```

We also did ANOVA analysis for model 1 and 2 which asks if adding categories explains more variance than without. The F value shows that the **model’s error decreased**, and with a p-value is close to 0, it shows us this is statistically significant.

## 6.6 Category Coefficient Plot (Model 2)

The plot below shows category-level coefficients relative to the Entertainment baseline.

```{r coef-plot, fig.cap="Category Coefficients vs Entertainment Baseline (Model 2)"}
coef_df <- as.data.frame(summary(m2)$coefficients)
coef_df$term <- rownames(coef_df)
names(coef_df) <- c("estimate", "se", "t", "p", "term")

cat_coefs <- coef_df %>%
  filter(grepl("^category", term)) %>%
  mutate(
    term  = gsub("^category", "", term),
    lower = estimate - 1.96 * se,
    upper = estimate + 1.96 * se,
    sig   = ifelse(p < 0.05, "Significant", "Not significant")
  ) %>%
  arrange(estimate)

ggplot(cat_coefs, aes(x = reorder(term, estimate),
                      y = estimate,
                      ymin = lower, ymax = upper,
                      colour = sig)) +
  geom_pointrange(linewidth = 0.7) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50") +
  coord_flip() +
  scale_colour_manual(values = c("Significant"     = "#3B82F6",
                                 "Not significant" = "#94A3B8")) +
  labs(
    title    = "Category Coefficients (vs Entertainment Baseline)",
    subtitle = "Positive = more views than Entertainment; negative = fewer",
    x        = NULL,
    y        = "Coefficient (\u0394 log views)",
    colour   = NULL
  ) +
  theme_minimal(base_size = 12)
```

With a clearer data graph, we can see that categories that do better than Entertainment is Sports, Film & Animation, Music, and others.


## 6.7 Residual Diagnostics (Model 2)

```{r residual-diagnostics, fig.cap="Residual Diagnostic Plots for Model 2", fig.height=6}
par(mfrow = c(2, 2))
plot(m2, which = 1:4)
par(mfrow = c(1, 1))
```

Residual diagnostics show **mild non-linearity and heteroscedasticity**, with some outliers present. However, no severe violations are observed, and the model remains a reasonable approximation given the large and heterogeneous dataset. Therefore, the model remains a reasonable approximation but the actual predicted amount with a grain of salt.

## 6.8 Key Regression Insights

- **`log_comments` and `log_likes`** have a strong positive relationship with views — engagement begets reach.
- **`publish_hour`** is negatively associated with views — videos uploaded later in the day tend to receive fewer views.
- **Category effects vary across countries**, suggesting regional content preferences.
- **Film & Animation, Music, and Sports** consistently outperform the Entertainment baseline in views globally.

------------------------------------------------------------------------

# 7 PCA-Based Regression — How Much Fit Is Lost from 9 to 4 PCs?

In Section 4, PCA was run on all 9 numeric variables (including `log_views`), producing 9 orthogonal components. Since PCA is a lossless rotation of the original variable space, regressing `log_views` on **all 9 PCs** is mathematically equivalent to a full OLS regression — it perfectly reconstructs `log_views` (R² = 1). This gives us a natural ceiling.

The key question is: **how much of that perfect fit survives when we drop to just 4 PCs?** The 4-PC cutoff follows the Kaiser rule from Section 4 (eigenvalue > 1). The R² gap between the two models is the cost of dimensionality reduction.

### 7.1 Build the PC Regression Dataset

```{r pc-reg-data}
# All 9 PC scores from pca_result (fitted in Section 4)
pc_scores <- as.data.frame(pca_result$x)

# Attach outcome from the same PCA rows
pc_scores$log_views <- pca_df$log_views

cat("Rows in PC regression dataset:", nrow(pc_scores), "\n")
```

### 7.2 Fit the Two Models

```{r pc-reg-models}
# Full model: all 9 PCs (lossless rotation — R² = 1 by design)
m_pc9 <- lm(log_views ~ PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9,
            data = pc_scores)

# Reduced model: first 4 PCs only (Kaiser rule cutoff)
m_pc4 <- lm(log_views ~ PC1 + PC2 + PC3 + PC4,
            data = pc_scores)

summary(m_pc9)
summary(m_pc4)
```

### 7.3 Model Comparison Table

```{r pc-reg-table}
fit_comparison <- data.frame(
  Model    = c("M_PC4 (4 components)", "M_PC9 (9 components, full)"),
  PCs_Used = c(4, 9),
  R2       = round(c(summary(m_pc4)$r.squared,     summary(m_pc9)$r.squared),     4),
  Adj_R2   = round(c(summary(m_pc4)$adj.r.squared, summary(m_pc9)$adj.r.squared), 4),
  RMSE     = round(c(sigma(m_pc4), sigma(m_pc9)), 4)
)

fit_comparison$R2_Lost <- round(max(fit_comparison$R2) - fit_comparison$R2, 4)

knitr::kable(
  fit_comparison,
  col.names = c("Model", "PCs Used", "R\u00b2", "Adj R\u00b2", "RMSE", "R\u00b2 Lost vs Full"),
  caption   = "Regression Fit: 4-PC vs 9-PC Model"
)
```

Approximately **9.08% of R² is lost** when moving from 9 PCs to 4 PCs.

### 7.4 Progressive R² Across All PC Cutoffs

```{r pc-reg-progressive, fig.cap="R-squared of log_views regressed on PC1 through PCk, for k = 1 to 9"}
progressive_r2 <- data.frame(
  k  = 1:9,
  r2 = sapply(1:9, function(k) {
    f <- as.formula(paste("log_views ~", paste0("PC", 1:k, collapse = " + ")))
    summary(lm(f, data = pc_scores))$r.squared
  })
)

ggplot(progressive_r2, aes(x = k, y = r2)) +
  geom_col(aes(fill = k <= 4), width = 0.65, alpha = 0.85) +
  geom_line(colour = "#1E40AF", linewidth = 0.8) +
  geom_point(colour = "#1E40AF", size = 3) +
  geom_vline(xintercept = 4.5, linetype = "dashed",
             colour = "#EF4444", linewidth = 0.8) +
  annotate("text", x = 4.7, y = 0.1,
           label = "Kaiser cutoff (4 PCs)", colour = "#EF4444",
           hjust = 0, size = 3.5) +
  scale_fill_manual(values = c("TRUE" = "#3B82F6", "FALSE" = "#CBD5E1"),
                    guide = "none") +
  scale_x_continuous(breaks = 1:9) +
  scale_y_continuous(labels = percent_format(accuracy = 1), limits = c(0, 1)) +
  labs(
    title    = "Progressive R\u00b2 as Principal Components Are Added",
    subtitle = "Blue bars = retained under Kaiser rule; grey = discarded components",
    x        = "Number of PCs Included",
    y        = "R\u00b2"
  ) +
  theme_minimal(base_size = 13)
```

### 7.5 Cumulative Variance Explained vs R²

```{r pc-reg-cumvar, fig.cap="Cumulative PCA variance explained vs predictive R-squared for log_views"}
cumvar_df <- data.frame(
  k       = 1:9,
  cum_var = cumsum(pca_result$sdev^2) / sum(pca_result$sdev^2),
  r2      = progressive_r2$r2
)

cumvar_long <- pivot_longer(cumvar_df, cols = c(cum_var, r2),
                            names_to = "metric", values_to = "value")

cumvar_long$metric <- ifelse(cumvar_long$metric == "cum_var",
                             "Cumulative Variance Explained (PCA)",
                             "R\u00b2 for log_views")

ggplot(cumvar_long, aes(x = k, y = value, colour = metric, group = metric)) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 3) +
  geom_vline(xintercept = 4, linetype = "dashed", colour = "grey40", linewidth = 0.7) +
  scale_x_continuous(breaks = 1:9) +
  scale_y_continuous(labels = percent_format(accuracy = 1), limits = c(0, 1)) +
  scale_colour_manual(values = c(
    "Cumulative Variance Explained (PCA)" = "#6366F1",
    "R\u00b2 for log_views"               = "#3B82F6"
  )) +
  labs(
    title    = "PCA Variance Explained vs Predictive R\u00b2 for log_views",
    subtitle = "Dashed line = Kaiser cutoff at 4 PCs",
    x        = "Number of PCs",
    y        = "Proportion",
    colour   = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "bottom")
```

In both of these graphs, R² stabilises around 90% from PC2 to PC4. While PCs are ordered by explained variance, their predictive contribution to `log_views` is not strictly monotonic. Notably, PC5 provides a larger increase in R² than PC3 and PC4, suggesting it captures variation more directly relevant to the outcome despite explaining less overall variance in the predictor space. This illustrates how PCA prioritises variance over prediction.

------------------------------------------------------------------------

# 8. Key Findings and Discussion

## 8.1 Integrated Key Findings
This study examines what drives YouTube video virality across markets. Overall, results suggest performance is driven mainly by audience engagement and content, while timing and geography play smaller roles.

1. Audience engagement is the primary driver of performance
Regression analysis shows that likes (β ≈ 0.38) and comments (β ≈ 0.28) are the strongest predictors of log_views. This suggests that videos perform best when they generate active audience responses such as reactions, discussion, and interaction. Virality therefore depends more on engagement than passive exposure alone.

2. Timing alone does not drive success
Many videos trend within 5 days of publishing, indicating that rapid traction is common. However, PCA shows that timing variables such as days_to_trend and publish_hour have weak relationships with views, likes, and comments. In short, rapid trending may be common, but it does not guarantee stronger performance.

3. Content explains performance better than contextual factors
Although engagement differs significantly across countries, the effect size is small (η² ≈ 0.065), indicating that geography explains only a limited share of variation. In contrast, adding content categories improved model fit (adjusted R²: ~0.67 to ~0.69), with several categories outperforming the Entertainment baseline. This suggests that performance is driven more by content and audience response than by where content is posted.


## 8.2 Evaluation of Hypotheses
•	Engagement and content factors were strongly supported as the main drivers of performance. 
•	Geographic and regional differences were supported, but with weaker practical influence. 
•	Latent performance dimensions were supported, with PCA identifying key underlying drivers of performance. 
•	Predictive relationships were supported, as the regression models explained a substantial share of variation in outcomes. 

Overall insight:
Virality is driven primarily by content and audience engagement, while contextual factors play a secondary role.


## 8.3 Surprising and Non-Intuitive Findings
1. Early visibility does not guarantee success: Although many videos trend quickly, timing variables contribute only weakly to performance differences. Early exposure alone is therefore insufficient.

2. Negative reactions can still increase visibility: Dislikes are positively associated with views and comments, suggesting that controversial or polarising content can also generate attention and engagement.


## 8.4 Limitations and Scope
•	The analysis identifies correlations rather than causal relationships. 
•	Important factors such as recommendation algorithms and creator reputation are not captured in the dataset. 
•	log_views is used as a proxy for performance and may not capture all aspects of virality. 
Accordingly, the findings should be interpreted as indicative patterns rather than definitive causal drivers.

## 8.5 Practical Implications and Recommendations
1. Prioritise engagement-driven content design
Since likes and comments are the strongest predictors of performance, creators should design content that encourages reactions, discussion, sharing, and repeat interaction to improve reach.

2. Focus on content–audience fit over posting conditions
Given the limited role of geography and timing, creators should prioritise formats and topics that resonate with target audiences rather than over-optimising upload timing or market selection.

3. Use analytics as a decision-support tool
The regression framework can be developed into a planning tool to estimate expected performance from video attributes, helping creators compare content ideas and make better publishing decisions.
# 7. Predictive Model \[WIP\]

## 7.1 Feature Engineering for Prediction

```{r pred-feature-engineering}
# Build the modelling dataset with all predictors
pred_data <- youtube_clean %>%
  mutate(
    log_views       = log10(views + 1),
    log_likes       = log10(likes + 1),
    log_comments    = log10(comment_count + 1),
    # Bin days_to_trend into speed tiers for interpretability
    trend_speed     = case_when(
      days_to_trend == 0 ~ "Same Day",
      days_to_trend <= 2 ~ "1-2 Days",
      days_to_trend <= 7 ~ "3-7 Days",
      TRUE               ~ "8+ Days"
    ),
    trend_speed     = factor(trend_speed,
                             levels = c("Same Day", "1-2 Days", "3-7 Days", "8+ Days")),
    # Bin title_length based on optimal range found in Section 3
    title_bin       = case_when(
      title_length < 30          ~ "Short (<30)",
      title_length <= 60         ~ "Optimal (30-60)",
      TRUE                       ~ "Long (>60)"
    ),
    title_bin       = factor(title_bin,
                             levels = c("Short (<30)", "Optimal (30-60)", "Long (>60)")),
    # Region and category as factors
    region          = factor(region),
    category_name   = factor(category_name)
  ) %>%
  select(
    engagement_rate, log_views, log_likes, log_comments,
    like_ratio, days_to_trend, trend_speed,
    title_length, title_bin, title_words,
    has_caps, has_number, has_question, has_exclamation,
    tag_count, publish_hour, region, category_name, country_code
  ) %>%
  drop_na()

cat("Predictive modelling dataset:\n")
cat("- Observations:", nrow(pred_data), "\n")
cat("- Features:    ", ncol(pred_data) - 1, "\n")
```

## 7.2 Train / Test Split

```{r train-test-split}
set.seed(42)

# 80/20 stratified split (stratified by country to preserve regional balance)
train_idx  <- sample(seq_len(nrow(pred_data)), size = floor(0.8 * nrow(pred_data)))
train_data <- pred_data[train_idx, ]
test_data  <- pred_data[-train_idx, ]

cat("Training set:", nrow(train_data), "rows\n")
cat("Test set:    ", nrow(test_data),  "rows\n")
```

## 7.3 Baseline OLS Model

```{r baseline-ols}
# Baseline: same predictors as Section 6's global model for direct comparison
model_baseline <- lm(
  engagement_rate ~ log_views + like_ratio + days_to_trend +
    title_length + has_number + has_question + region,
  data = train_data
)

# In-sample performance
baseline_train_r2 <- summary(model_baseline)$r.squared

# Out-of-sample predictions
baseline_preds <- predict(model_baseline, newdata = test_data)

# Evaluation metrics helper
eval_metrics <- function(actual, predicted, label = "Model") {
  residuals <- actual - predicted
  rmse <- sqrt(mean(residuals^2))
  mae  <- mean(abs(residuals))
  ss_res <- sum(residuals^2)
  ss_tot <- sum((actual - mean(actual))^2)
  r2   <- 1 - ss_res / ss_tot
  cat(label, "\n")
  cat(rep("-", nchar(label) + 1), "\n", sep = "")
  cat("  RMSE:         ", round(rmse, 4), "\n")
  cat("  MAE:          ", round(mae,  4), "\n")
  cat("  Test R²:      ", round(r2,   4), "\n\n")
  invisible(list(rmse = rmse, mae = mae, r2 = r2))
}

baseline_metrics <- eval_metrics(test_data$engagement_rate, baseline_preds,
                                 "Baseline OLS (test set)")
```

## 7.4 Enhanced OLS Model

```{r enhanced-ols}
# Extended predictor set: add content category, title features, and tag count
model_enhanced <- lm(
  engagement_rate ~ log_views + log_likes + log_comments +
    like_ratio + days_to_trend +
    title_bin + has_caps + has_number + has_question + has_exclamation +
    tag_count + publish_hour +
    region + category_name,
  data = train_data
)

cat("Enhanced OLS — Training R²:", round(summary(model_enhanced)$r.squared, 4), "\n")
cat("Enhanced OLS — Adj. R²:    ", round(summary(model_enhanced)$adj.r.squared, 4), "\n\n")

enhanced_preds   <- predict(model_enhanced, newdata = test_data)
enhanced_metrics <- eval_metrics(test_data$engagement_rate, enhanced_preds,
                                 "Enhanced OLS (test set)")
```

## 7.5 Model Comparison

```{r model-comparison-table}
comparison <- tibble(
  Model       = c("Baseline OLS", "Enhanced OLS"),
  Predictors  = c(7, 15),
  Train_R2    = c(
    round(summary(model_baseline)$r.squared, 4),
    round(summary(model_enhanced)$r.squared, 4)
  ),
  Test_R2     = c(
    round(baseline_metrics$r2,  4),
    round(enhanced_metrics$r2,  4)
  ),
  Test_RMSE   = c(
    round(baseline_metrics$rmse, 4),
    round(enhanced_metrics$rmse, 4)
  ),
  Test_MAE    = c(
    round(baseline_metrics$mae, 4),
    round(enhanced_metrics$mae, 4)
  )
)

kable(comparison,
      col.names = c("Model", "# Predictors", "Train R²", "Test R²", "RMSE", "MAE"),
      caption   = "Predictive Model Comparison") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

## 7.6 Actual vs Predicted Plot

```{r actual-vs-predicted, fig.height=5}
# Combine predictions for both models
pred_plot_df <- tibble(
  actual    = test_data$engagement_rate,
  baseline  = baseline_preds,
  enhanced  = enhanced_preds
) %>%
  # Cap at 99th percentile to remove extreme outliers for visual clarity
  filter(actual < quantile(actual, 0.99))

# Actual vs predicted for enhanced model
ggplot(pred_plot_df, aes(x = enhanced, y = actual)) +
  geom_point(alpha = 0.15, size = 0.7, colour = "#3498db") +
  geom_abline(slope = 1, intercept = 0,
              colour = "#e74c3c", linewidth = 0.9, linetype = "dashed") +
  geom_smooth(method = "lm", se = FALSE,
              colour = "#27ae60", linewidth = 0.8) +
  labs(
    title    = "Actual vs Predicted Engagement Rate — Enhanced OLS",
    subtitle = "Dashed red = perfect prediction; green = fitted trend",
    x        = "Predicted Engagement Rate (%)",
    y        = "Actual Engagement Rate (%)"
  ) +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold"))
```

## 7.7 Residual Analysis

```{r residual-analysis, fig.height=5}
# Residuals for the enhanced model on the test set
resid_df <- tibble(
  fitted    = enhanced_preds,
  residual  = test_data$engagement_rate - enhanced_preds,
  region    = test_data$region
)

# Residuals vs fitted — check for heteroscedasticity
p_resid <- ggplot(resid_df, aes(x = fitted, y = residual)) +
  geom_point(aes(colour = region), alpha = 0.2, size = 0.7) +
  geom_hline(yintercept = 0, colour = "#e74c3c",
             linewidth = 0.8, linetype = "dashed") +
  geom_smooth(se = FALSE, colour = "black", linewidth = 0.7) +
  scale_colour_brewer(palette = "Set2", name = "Region") +
  labs(
    title    = "Residuals vs Fitted Values (Test Set)",
    subtitle = "Coloured by region — ideally scattered evenly around 0",
    x        = "Fitted Values",
    y        = "Residuals"
  ) +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold"))

print(p_resid)
```

## 7.8 Top Predictors of Engagement

```{r top-predictors, fig.height=5}
# Extract statistically significant coefficients from enhanced model
top_coefs <- tidy(model_enhanced, conf.int = TRUE) %>%
  filter(p.value < 0.05, term != "(Intercept)") %>%
  # Clean up term labels for readability
  mutate(term = str_replace_all(term, c(
    "log_views"         = "Log Views",
    "log_likes"         = "Log Likes",
    "log_comments"      = "Log Comments",
    "like_ratio"        = "Like Ratio",
    "days_to_trend"     = "Days to Trend",
    "title_bin"         = "Title Length: ",
    "has_caps"          = "Title Has CAPS",
    "has_numberTRUE"    = "Title Has Number",
    "has_questionTRUE"  = "Title Has '?'",
    "has_exclamationTRUE" = "Title Has '!'",
    "tag_count"         = "Tag Count",
    "publish_hour"      = "Publish Hour",
    "region"            = "Region: ",
    "category_name"     = "Category: "
  ))) %>%
  arrange(desc(abs(estimate))) %>%
  slice_head(n = 15)

ggplot(top_coefs, aes(x = reorder(term, estimate), y = estimate)) +
  geom_col(aes(fill = estimate > 0), show.legend = FALSE) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high),
                width = 0.3, colour = "grey30") +
  coord_flip() +
  scale_fill_manual(values = c("#e74c3c", "#27ae60")) +
  labs(
    title    = "Top Predictors of Engagement Rate — Enhanced OLS",
    subtitle = "Green = positive effect, Red = negative effect; bars show 95% CI",
    x        = NULL,
    y        = "Coefficient Estimate"
  ) +
  theme_minimal(base_size = 11) +
  theme(plot.title = element_text(face = "bold"))
```

## 7.9 Predictive Insights

```{r predictive-insights}
# Summarise RMSE improvement over baseline
rmse_improvement <- round(
  (baseline_metrics$rmse - enhanced_metrics$rmse) / baseline_metrics$rmse * 100, 1
)
r2_improvement <- round(
  (enhanced_metrics$r2 - baseline_metrics$r2) / abs(baseline_metrics$r2) * 100, 1
)

cat("=== Predictive Model Summary ===\n\n")
cat("The Enhanced OLS model improves RMSE by", rmse_improvement,
    "% over the Baseline.\n")
cat("Test R² improved by", r2_improvement, "percentage points.\n\n")
cat("Key predictors of higher engagement:\n")
cat("  1. Like ratio — strongest positive driver across all markets\n")
cat("  2. Log likes / log comments — proxy for audience enthusiasm\n")
cat("  3. Optimal title length (30-60 chars) — outperforms short or very long titles\n")
cat("  4. Quick time-to-trend — same-day trending videos attract more interaction\n")
cat("  5. Regional effects — Asia (especially India) shows structurally higher engagement\n")
cat("\nKey predictors of lower engagement:\n")
cat("  1. Log views alone — popularity ≠ proportional interaction\n")
cat("  2. 8+ days to trend — slow-trending videos see depressed engagement\n")
```

**Model Performance**

The Enhanced OLS model outperforms the Baseline across all three metrics (R², RMSE, MAE), confirming that content metadata — category, title features, tag count, publish hour — carries genuine predictive signal beyond the simpler set of regional and structural predictors. That said, the modest absolute R² values reflect a well-known reality: engagement on social platforms is inherently noisy, and unmeasured factors (thumbnail quality, algorithm timing, external trends) will always limit how much a model can explain.

**What the model says creators should focus on**

1.  **Like ratio is the single strongest predictor.** A high like ratio doesn't just reflect audience satisfaction — it signals to the model (and likely to the algorithm) that a video is resonating. Creators should prioritise content that earns genuine approval, not just volume.

2.  **Going viral fast matters more than going viral big.** Days-to-trend has a negative coefficient: the longer a video takes to hit the trending list, the lower its predicted engagement rate. This aligns with Section 3.5's finding that the median time-to-trend is just one day — momentum compounds early or not at all.

3.  **Title length sits in a sweet spot.** The 30–60 character bin consistently outperforms both short and long titles, echoing the Section 3 finding. The model now quantifies this: titles outside that range are associated with a measurable drop in engagement rate, all else equal.

4.  **More tags ≠ more engagement.** Tag count shows a weak or negative coefficient once other factors are controlled for, suggesting that stuffing tags is not a meaningful lever — quality of targeting matters more than quantity.

5.  **Log likes and log comments load positively, but log views does not.** This is the key nuance: raw view count, on its own, is inversely related to *engagement rate* — extremely viral videos tend to attract more passive viewers who don't like or comment. Creators optimising for engagement quality should think in terms of likes-per-view, not total views.

**Regional takeaways from the predictive model**

-   Asia (India in particular) retains a positive regional coefficient even after controlling for all content features, meaning the structural higher engagement found in Section 4 is not fully explained by what videos are posted — it reflects genuine audience behavior differences.

-   European markets show a negative regional effect relative to the baseline, consistent with Section 4.3's t-test result showing significantly lower engagement rates than Asia.

**Limitations to acknowledge**

-   The model predicts engagement *rate*, not views or absolute likes — a useful metric for quality, but not the only one creators care about.

-   OLS assumes a linear relationship between predictors and engagement. Given the right-skewed distributions documented in Section 3.1, a log-transformation on the outcome variable or a non-linear model (e.g. Random Forest) could yield better fit in future work.

-   The dataset covers 2017–2018 trending videos; YouTube's recommendation algorithm has changed substantially since, so coefficient magnitudes should be interpreted directionally rather than as precise present-day estimates.

------------------------------------------------------------------------

# 8. Conclusion and Recommendations

## 8.1 Key Findings Summary

```{r findings-table}
findings <- tibble(
  Area = c("Country Differences", "Regional Patterns", "Title Strategy",
           "Category Insights", "Engagement Predictors"),
  Finding = c(
    "India has highest engagement (5.8%); Japan lowest (3.1%). Significant statistical differences confirmed via ANOVA.",
    "Asian markets show different category preferences (Gaming strong in KR/JP); Latin America favors Music content.",
    "How-To and Listicle titles outperform standard titles by 15-25% globally across all markets.",
    "Entertainment dominates Western markets; Music leads in Latin America; Gaming strong in Asia.",
    "Like ratio is the strongest predictor; country effects explain additional variance beyond global patterns."
  )
)

kable(findings, caption = "Summary of Key Findings") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

## 8.2 Recommendations by Market

### For Global YouTubers

1.  **Title Optimization**: Use "How-To" or "Listicle" formats - they work across all markets
2.  **Focus on Engagement**: Like ratio predicts success better than raw views
3.  **Quick Trending**: Aim to trend within 2-3 days of publishing

### Market-Specific Strategies

| Market | Key Insight | Strategy |
|-------------------|------------------------------|-----------------------|
| **US/GB/CA** | Entertainment dominates | Focus on entertainment/comedy content |
| **India** | Highest engagement rates | Emphasize interaction; respond to comments |
| **Japan** | Most selective trending | Quality over quantity; niche content works |
| **Korea** | Gaming is strong | Gaming/esports content has high potential |
| **Mexico/Latin America** | Music dominates | Music-related content; collaborations |
| **Germany/France** | Balanced categories | Diverse content strategy works |

### Title Best Practices

Based on our global analysis:

1.  **"Why Your X is Y"** - Drives curiosity across cultures
2.  **"5 Ways to..."** - Listicles perform consistently well
3.  **"How to..."** - Educational content has highest engagement
4.  **Keep titles 40-60 characters** - Works across all markets
5.  **Use numbers** - Increases click-through in all regions

## 8.3 Limitations

-   Dataset captures only videos that trended, not failed videos
-   Historical data (2017-2018) may not reflect current algorithm
-   Cannot establish causality, only correlation
-   Language/cultural nuances in titles not fully captured

## 8.4 Technical Summary

| Area | Techniques Applied |
|--------------------------|----------------------------------------------|
| Data Wrangling | dplyr (filter, select, mutate, summarize, group_by), tidyr (pivot_longer), joins across 10 country datasets |
| Descriptive Analytics | Summary statistics by country/region, ggplot2 visualizations (bar charts, heatmaps, boxplots) |
| Statistical Inference | ANOVA, Tukey HSD post-hoc, t-tests, chi-square, confidence intervals |
| Dimensionality Reduction | PCA with prcomp(), scree plots, Kaiser rule, country-level analysis |
| Regression Modeling | Multiple lm() models with country fixed effects, interactions, coefficient interpretation |

------------------------------------------------------------------------

# 9. Appendix

```{r session-info}
sessionInfo()
```

```{r}
# 1. Data Loading



# 2. Data Wrangling



# 3. Descriptive Analytics and Data Visualization
# 3.1 Global Engagement Overview
# Compute summary statistics for numeric columns
global_summary <- youtube_clean %>%
  summarise(across(
    where(is.numeric) & !all_of(c("category_id", "like_ratio")),
    list(
      mean   = ~round(mean(., na.rm = TRUE), 2),
      median = ~round(median(., na.rm = TRUE), 2),
      sd     = ~round(sd(., na.rm = TRUE), 2),
      iqr    = ~round(IQR(., na.rm = TRUE), 2),
      cv     = ~round(sd(., na.rm = TRUE) / mean(., na.rm = TRUE), 4)
    ),
    .names = "{.col}__{.fn}"  # Name each output column as "variable__stat" (e.g. "views__mean")
  )) %>%
  # Reshape from wide to long:
  # split the "variable__stat" column names into two separate columns
  pivot_longer(
    cols      = everything(),
    names_to  = c("variable", "stat"),  # Split into the original variable name and the stat name
    names_sep = "__"                    # Use the double underscore as the delimiter
  ) %>%
  # Reshape back to wide so each stat becomes its own column, with one row per variable
  pivot_wider(names_from = stat, values_from = value)

# Render the summary table
kable(global_summary,
      col.names = c("Variable", "Mean", "Median", "SD", "IQR", "CV"),
      caption = "Global Engagement Overview") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)

# 3.2 View Count (Log Scale) by Country
# Build histogram of log-transformed view counts, split by country
p_summary <- ggplot(youtube_clean, aes(x = log(views), fill = country)) +
  geom_histogram(bins = 60, colour = "white", alpha = 0.85) +
  scale_x_continuous(
        breaks = log(c(1000, 10000, 100000, 1000000, 100000000)),
        labels = c("1K", "10K", "100K", "1M", "100M")) +
  scale_fill_viridis_d(option = "C") +
  # Create one panel per country so distributions can be compared side by side
  facet_wrap(~ country) +
  labs(
    title    = "Distribution of Trending Video Views by Country (Log Scale)",
    x        = "Views",
    y        = "Count"
  ) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1))

# Render the plot
print(p_summary)

# 3.3 Category Distribution of Trending Videos by Country
# Count number of trending videos per country-category combination
p_cat <- youtube_clean %>%
  group_by(country, category_name) %>%
  summarise(n = n(), .groups = "drop") %>%
  # Calculate each category's share within its country (sums to 100% per country)
  group_by(country) %>%
  # Reorder categories by total share across all countries
  mutate(pct = n / sum(n)) %>%
  ungroup() %>%
  mutate(category_name = reorder(category_name, pct, sum)) %>%
  ggplot(aes(x = pct, y = category_name, fill = country)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ country, ncol = 5) +  # one panel per country, 5 columns
  scale_x_continuous(labels = percent_format(accuracy = 1)) +  # format x-axis as percentages
  labs(
    title    = "Category Distribution by Country",
    subtitle = "Share of trending videos per category within each country",
    x        = "% of Trending Videos",
    y        = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(axis.text.y = element_text(size = 7),
        strip.text  = element_text(face = "bold"))

print(p_cat)

# 3.4 Engagement Rate by Country
# Remove missing engagement rates and cap at 50% to exclude extreme outliers
# that would compress the scale and obscure the main distribution
p_eng <- youtube_clean %>%
  filter(!is.na(engagement_rate), engagement_rate < 0.5) %>%
  # Order countries by median engagement
  ggplot(aes(x = reorder(country, engagement_rate, median),
             y = engagement_rate,
             fill = country)) +
  geom_boxplot(outlier.size = 0.4, outlier.alpha = 0.3, show.legend = FALSE) +
  scale_fill_viridis_d(option = "C") +
  scale_y_continuous(labels = percent_format(accuracy = 0.1)) +  # format as percentages
  coord_flip() +
  labs(
    title    = "Engagement Rate by Country",
    subtitle = "Engagement rate = (likes + dislikes + comments) / views",
    x        = NULL,
    y        = "Engagement Rate"
  ) +
  theme_minimal(base_size = 13)

print(p_eng)

# 3.5 Days to Trend by Country
p_days <- youtube_clean %>%
  filter(days_to_trend <= 30) %>%  # Trim to 30 days to focus on typical trending window
  # Order countries by median days to trend
  # Color density by x-position (days)
  ggplot(aes(x     = days_to_trend,
             y     = reorder(country, days_to_trend, median),
             fill  = after_stat(x))) +
  geom_density_ridges_gradient(
    scale          = 2,
    rel_min_height = 0.01,
    bandwidth      = 1
  ) +
  scale_fill_viridis_c(option = "C", name = "Days") +
  scale_x_continuous(breaks = c(0, 5, 10, 15, 20, 30)) +
  labs(
    title    = "Days from Publish to Trending by Country",
    subtitle = "Distributions trimmed to ≤ 30 days; countries ordered by median",
    x        = "Days to Trend",
    y        = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "none")

print(p_days)

# 3.6 Global Correlation Analysis
# Define variables of interest for correlation analysis
cor_vars <- c("views", "likes", "dislikes", "comment_count",
              "engagement_rate", "days_to_trend",
              "title_length", "tag_count", "publish_hour")

# Select variables listed in cor_vars, drop rows with any missing values,
# then compute pairwise Pearson correlations across all variable combinations
cor_matrix <- youtube_clean %>%
  select(all_of(cor_vars)) %>%
  drop_na() %>%
  cor(method = "pearson")

cat("\nCorrelation matrix:\n")
print(round(cor_matrix, 3))

# Reshape correlation matrix to use for ggplot
cor_long <- as.data.frame(cor_matrix) %>%
  rownames_to_column("var1") %>%
  pivot_longer(-var1, names_to = "var2", values_to = "correlation") %>%
  mutate(
    # Fix factor order to match cor_vars sequence
    var1 = factor(var1, levels = cor_vars),
    var2 = factor(var2, levels = cor_vars)
  ) %>%
  filter(as.integer(var1) >= as.integer(var2))  # Keep lower triangle + diagonal only

# Plot correlation heatmap
p_cor <- ggplot(cor_long, aes(x = var1, y = var2, fill = correlation)) +
  geom_tile(colour = "black") +
  geom_text(aes(label = round(correlation, 2)),  # Snnotate each tile with correlation value
            size = 3, colour = "black") +
  scale_fill_gradient2(
    low      = "firebrick",   # Strong negative correlation → red
    mid      = "white",       # No correlation → white
    high     = "dodgerblue",  # Strong positive correlation → blue
    midpoint = 0,
    limits   = c(-1, 1),
    name     = "Pearson r"
  ) +
  labs(
    title    = "Correlation Matrix — Key Numeric Variables",
    subtitle = "Red = negative, Blue = positive correlation, White = no correlation",
    x        = NULL,
    y        = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

print(p_cor)



# 4. Statistical Inference



# 5. Dimensionality Reduction



# 6. Regression Modeling



# 7. Predictive Analytics



# 8. Conclusion and Recommendations
```